Clustering Beat-Chroma Patterns\\ in a Large Music Database

### 2010-03-24 Paper/Paper.tex
more updates to section 2

\subsection{Definitions}
A pattern, or a patch, is a beat-chroma pattern taken from an actual song.
A codeword (sometimes just code) is similar to a pattern, but is the result 
of the clustering
algorithm and most likely does not come from an actual song.
Distortion is our measure of distance between a pattern and a code
(or between two patterns, see equation \ref{eq:dist} in Section 
\ref{sec:experiments}).

% COMMENTING PREVIOUS WORK
\iffalse
\subsection{Previous Work}\label{sec:prevwork}
Our work can be seen as an extension of the \textit{Shingles} idea
described by Casey and Slaney \cite{Casey2006,Casey2007,Casey2008}.
The two main differences are the clustering framework and the use of
beats and bars to select the position of the \textit{Shingles}.

@@

\section{Approach}\label{sec:approach}
In this section we present our use of the Echo Nest features and the
datasets we use for training and testing our model.

\subsection{Features}
We use the Echo Nest analyze API \cite{EchoNest} throughout this work.
The API gives us a chroma vector (length $12$) for every music event (called 
``segment'') for any song we upload to their platform. 
We also get an estimate of where the beats and the bars are. Beats span over
multiple segments, and bars span over multiple beats. 
We transform the chroma vectors per segment into per beat vectors using a 
simple average. We can then stack beat vectors for a number of bars 
(usually 1 or 2). 
We get a fixed-size patch by resampling the number of chroma vectors. Typically,
there are $4$ beats per bar (Figure \ref{fig:barsize}). 
Thus an appropriate pattern size for one bar
is $12$x$4$ or $12$x$8$.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.9\columnwidth]{bar_size_freq}
\end{center}
\caption{\small{
Frequency of the different bar size in the training data.
}}
\label{fig:barsize}
\end{figure}

Note that we do not claim that any of these informations (segments, beats, bars)
are perfectly accurate (\eg in \cite{Barrington2009a} authors obtain a better
song segmentation than Echonest). Practise showed us that they are reasonable, 
and the size of the data set should make up for the imperfections or noise.
We also believe that patches sizes based on a number of beats or bars are more
meaningful than an arbitrary time length. More on this in the experiments
(Subsection \ref{ssec:segment}).

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.6\columnwidth]{code}
\end{center}
\caption{\small{A typical codeword from a codebook of size $200$. 
Patterns represented
$2$ bars and the pattern length was set to $8$.
}}
\label{fig:code}
\end{figure}

%\section{Algorithm}\label{sec:algo}
In this section we present the creation of the harmonic patterns from
the Echo Nest data.  Then we explain the online vector quantization algorithm
used to learn a codebook of those patterns.

@@

We use the bar segmentation obtained from the analysis described in
the previous section to break a song into a collection of beat-chroma
``patches'', typically one or two bars in length. 
%
Because the bar length varies, depending on the meter of a particular
song (see Figure \ref{fig:barsize}), we resample each patch to a fixed
length of typically 4 to 8 beats.
% although the majority of the data contains bars that are 4 beats long

@@

>>>>

% COMMENTING PREVIOUS WORK AND DEFINITIONS
\iffalse
\subsection{Definitions}
A pattern, or a patch, is a beat-chroma pattern taken from an actual song.
A codeword (sometimes just code) is similar to a pattern, but is the result 
of the clustering
algorithm and most likely does not come from an actual song.
Distortion is our measure of distance between a pattern and a code
(or between two patterns, see equation \ref{eq:dist} in Section 
\ref{sec:experiments}).

\subsection{Previous Work}\label{sec:prevwork}
Our work can be seen as an extension of the \textit{Shingles} idea
described by Casey and Slaney \cite{Casey2006,Casey2007,Casey2008}.
The two main differences are the clustering framework and the use of
beats and bars to select the position of the \textit{Shingles}.

@@

\section{Approach}\label{sec:approach}
% In this section we review the creation of the beat-chroma patterns using
% the Echo Nest data.  Then we explain the online vector quantization algorithm
% used to learn a codebook of those patterns.

\subsection{Features}
The feature analysis used throughout this work is based on Echo Nest
analyze API \cite{EchoNest}.  
%
For any song uploaded to their platform this analysis returns a chroma
vector (length $12$) for every music event (called ``segment''), and a
segmentation of the song into beats and bars. Beats span multiple
segments, and bars span multiple beats.
%
Averaging the per-segment chroma over beat times results in a
beat-synchronous chroma feature representation similar to that used in
\cite{Ellis2007a}.

% \begin{figure}[htb]
% \begin{center}
% \includegraphics[width=.9\columnwidth]{bar_size_freq}
% \end{center}
% \caption{\small{
% Frequency of the different bar size in the training data.
% }}
% \label{fig:barsize}
% \end{figure}

Note that we do not claim that any of these informations (segments, beats, bars)
are perfectly accurate (\eg in \cite{Barrington2009a} authors obtain a better
song segmentation than Echonest). Practice showed us that they are reasonable, 
and the size of the data set should make up for the imperfections or noise.
We also believe that patches sizes based on a number of beats or bars are more
meaningful than an arbitrary time length. More on this in the experiments
(Subsection \ref{ssec:segment}).

@@

We use the bar segmentation obtained from the Echo Nest analysis to
break a song into a collection of beat-chroma ``patches'', typically
one or two bars in length.
%
Because the bar length varies, depending on the meter of a particular
song, we resample each patch to a fixed length of typically 4 beats
per bar.  Since the overwhelming majority (85\%) of our training data
consisted of bars that were 4 beats long, this resampling was not
often required.  Most of the remaining bars (14\%) were 3 beats in
length.  (FIXME: thierry, can you check these numbers, I'm making them
up based on the deletedfigure)
% although the majority of the data contains bars that are 4 beats long
The resulting patches consist of $12 \times 4$ or $12 \times 8$ matrices.

\begin{figure}[thb]
\begin{center}
\includegraphics[width=.6\columnwidth]{code}
\end{center}
\caption{\small{A typical codeword from a codebook of size $200$. 
Patterns represented
$2$ bars and the pattern length was set to $8$.
}}
\label{fig:code}
\end{figure}

@@

% We can then stack beat vectors for a number of bars (usually 1 or 2).
